HealpixNet: Efficient spherical Convolutional Neural Network with Healpix sampling for cosmological applications

### 2018-07-06 scnn.tex
modified abstract

Convolutional Neural Networks (CNNs) are now a cornerstone of the Deep Learning toolbox and have led to many breakthroughs in Artificial Intelligence. Despite their successes, they still are mostly developed and applied to regular Euclidean domains such as those supporting images, audio, or video. Unfortunately, in many applications, the data domain is not regular.
As a response to that fact, we are witnessing an increasing interest in the generalization of CNNs to graphs. Graphs are versatile data structures which can represent pairwise relationships between objects or act as a discrete representation of a continuous manifold.
In this paper we take the latter view and represent a sampled sphere as a graph, for which we define the convolution and pooling operations.
Spherical maps appear in many applications, such as the study of phenomena on the Earth for meteorologists or on the sky for cosmologists.
A popular format for spherical maps is Healpix, which uses an equal-area, isolatitude sampling. Hence, we present a spherical CNN for analysis of full and partial Healpix maps, which we call HealpixNet.
\nati{I am not that we should talk about pooling in the abstract. Another angle would be to say that HealPix sampling is well suited for spherical CNN with graphs as a) it is very regular => the graph capture well the spherical structure of the Manifold and b) pooling is natural.} 
While the convolution on arbitrary graphs has mostly been sorted out, a general pooling operation has not been discovered so far. 
As such, while HealpixNet is a special case of a graph CNN, it is tailored to the Healpix sampling of the sphere.
This approach is computationally much more efficient than using spherical harmonics to perform convolutions, an alternative that has been explored previously.
% The filters learned by the HealpixNet are restricted to being radial.
% Michaël: not necessary in the abstract?
We demonstrate the method on a classification problem of mass maps from two cosmological models. 
\nati{Should we specify what are the baselines? I think not.}
A standard benchmark for that problem is to train an SVM on the histograms of map pixel values. Another is to use the power spectral density instead.
HealpixNet achieves \todo{xx\%} better accuracy than those baselines, is more robust to noise, and is \todo{faster?}.
Code and examples are available at \todo{\url{https://github.com/xxx}}.
\end{abstract}

>>>>

Convolutional Neural Networks (CNNs) are now a cornerstone of the Deep Learning toolbox and have led to many breakthroughs in Artificial Intelligence.
These networks have so far been developed mostly for regular Euclidean domains such as those supporting images, audio, or video.
CNN-based methods are becoming increasingly popular in Cosmology.
Cosmological data often comes in a form of spherical maps, which makes the use of CNNs more complicated.
A popular format for spherical maps is Healpix, which uses an equal-area, isolatitude sampling.
We present a spherical CNN for analysis of full and partial Healpix maps, which we call HealpixNet.
We construct a spherical CNNs by representing the sphere as a graph.
Graphs are versatile data structures which can represent pairwise relationships between objects or act as a discrete representation of a continuous manifold.
For that graph, we define many of the standard CNN operations, such as convolution and pooling, with the filters restricted to being radial.
HealpixNet is a special case of a graph CNN, tailored to the Healpix sampling of the sphere.
This approach is computationally more efficient than using spherical harmonics to perform convolutions, an alternative that has been explored previously.
We demonstrate the method on a classification problem of mass maps from two cosmological models.
We compare the performance of the CNN with that of two baseline classifiers, based on the power spectrum and pixel density histogram.
We find the performance of HealpixNet to always better or equal to both of these baselines.
For high noise levels and for data covering only a smaller fraction of the sphere, HealpixNet achieves typically 10\% better classification accuracy than those baselines.
Finally, we visualise filters learned by the HealpixNet from the mass maps.
Code and examples are available at \todo{\url{https://github.com/xxx}}.

%As a response to that fact, we are witnessing an increasing interest in the generalization of CNNs to graphs.
% Graphs are versatile data structures which can represent pairwise relationships between objects or act as a discrete representation of a continuous manifold.
% In this paper we take the latter view and represent a sampled sphere as a graph, for which we define the convolution and pooling operations.
% Spherical maps appear in many applications, such as the study of phenomena on the Earth for meteorologists or on the sky for cosmologists.
% A popular format for spherical maps is Healpix, which uses an equal-area, isolatitude sampling.
% Hence, we present a spherical CNN for analysis of full and partial Healpix maps, which we call HealpixNet.
% \nati{I am not that we should talk about pooling in the abstract. Another angle would be to say that HealPix sampling is well suited for spherical CNN with graphs as a) it is very regular => the graph capture well the spherical structure of the Manifold and b) pooling is natural.}
% While the convolution on arbitrary graphs has mostly been sorted out, a general pooling operation has not been discovered so far.
% As such, while HealpixNet is a special case of a graph CNN, it is tailored to the Healpix sampling of the sphere.
% This approach is computationally much more efficient than using spherical harmonics to perform convolutions, an alternative that has been explored previously.
% The filters learned by the HealpixNet are restricted to being radial.
% Michaël: not necessary in the abstract?
% We demonstrate the method on a classification problem of mass maps from two cosmological models.
% \nati{Should we specify what are the baselines? I think not.}
% A standard benchmark for that problem is to train an SVM on the histograms of map pixel values. Another is to use the power spectral density instead.
% HealpixNet achieves \todo{xx\%} better accuracy than those baselines, is more robust to noise, and is \todo{faster?}.
% Code and examples are available at \todo{\url{https://github.com/xxx}}.
\end{abstract}
